NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration

### 2022-10-12 main.tex
Update text

@@

\subsection{Water diffusion anisotropy in membrane systems}
The anisotropic water diffusion through and along of membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}, and in MRI imaging~\cite{??}. Therefore, dependence of water permeation through membranes and translational diffusion at membrane surfaces on membrane properties has been studied with different methods, yet consensus 
%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
has not been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}. MD have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{??}. Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.

To this end, we first calculated the water permeability through membranes from all the simulations in the NMRlipids databank. Resulting permeabilities are shown as a function of temperature, membrane thickness, area per lipid, and acyl chain order in Figs.~\ref{fig:permeability} B-E. The values for water permeabilities from simulations vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}. To analyze how water permeability through membranes depend on these properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range. These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots. As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\label{fig:permeability} B. Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order. Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai08,frallicciardi22}. Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence for more loosely packed membranes. 
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
%, although weak decrease for the latter two may be visible.

@@

To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water translational diffusion along the membrane surface from all simulations in the NMRlipids databank. The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}. 
%The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
%As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
Simulation results are close to the experimental values at low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) at high hydration levels. 
This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{??}. The water translational diffusion incereased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information. In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{??}. The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}. Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} F and G.
This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig. \ref{fig:diffusionSI} A and C).

@@

@@

Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance. In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions. This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist. Emerging applications of machine learning are increasing the impact of such databanks. For example, the existing Protein Databank (PDB)~\cite{??} containing experimentally determined protein structures with programmatic access has enabled the development of machine learning based tools building on the data collected in the databank over the years~\cite{??}. NMRlipids and other databanks with open programmatic access have potential to lead similar unforeseen applications in the future.

@@

\subsection{Structure of the databank}

\begin{figure}
  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
  \caption{Structure of the NMRlipids databank. Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available. The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes). Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line. }\label{DatabankStructure}
\end{figure}

Structure of the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure}. The required input information to create an entry into the NMRlipids databank are listed in table~\ref{tab:READMEkeys}. While the raw simulation data is not directly stored in the NMRlipids databank, permanent links from where the raw data can be accessed have to be given and are then stored in the README.yaml files at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}. These files contain all the essential information listed in table~\ref{tab:READMEkeys} on each simulation entry that are needed for further applications. The raw MD simulation data can locate in any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.

%A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data. In addition to this information, automatically extracted information about the simulation are stored in the README.yaml files (see table~\ref{tab:READMEkeys}). A README.yaml file for each simulation is stored to a folder named according to the hashes of original trajectory and topology file locating in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}. These README.yaml files give access to all the essential information on the simulations in the databank either directly, or pointing to a original file from where the information can be extracted.

\begin{table}[]
    \centering
    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
    \toprule
    key & description & type  \\
    \midrule
        %\hline
    DOI & DOI from where the raw data is found & user given (compulsory) \\
    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
    TRJ & Name of the trajectory file found from DOI & \\
    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds. \tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses. \tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
    DIR\_WRK & Temporary working directory in your local computer. \\
    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
    TYPEOFSYSTEM & Lipid bilayer or something else & \\
    \hline
    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
    SYSTEM & System description on free text format & \\
    SOFTWARE\_VERSION & Version of the used software & \\
    FF & Name of the used force field & \\
    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
    CPT & Name of the Gromacs checkpoint file. & \\
    LOG & Name of the Gromacs log file. & \\
    TOP & Name of the Gromacs top file. & \\
    GRO & Name of the Gromacs gro file. & \\
    \hline
    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
    TRJLENGTH & Lenght of the trajectory (ps). & \\
    TEMPERATURE & Temperature of the simulation. & \\
    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
    DATEOFRUNNIG & Date when added into the databank & \\
    EXPERIMENT & Potentially connected experimental data & \\
    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
    \end{tabular}
    \caption{Keys stored in the README.yaml files of simulations.}
    \label{tab:READMEkeys}
\end{table}

In order to evaluate the quality of simulations, sets of C-H bond order parameters from NMR and from factors from x-ray scattering are included in the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments}). The required information for an experimental dataset are listed in table~\ref{tab:READMEkeysEXP}.
A simulation is connected to a experimental data set
%C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script. Available experimental data is collected to  where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}. 
if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees. In such cases, a simulation and experimental data are paired by adding the path to the experimental data into the simulation README.yaml file. 
%If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.

\begin{table}[]
    \centering
    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
    \toprule
    key & description \\
    \midrule
    DOI & DOI of the publication related to the experimental data. \\
    TEMPERATURE & Temperature of the experiment. \\
    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
    COUNTER\_IONS & Type of counter ions if present.
\end{tabular}
    \caption{Keys stored in the README.yaml files of experiments.}
    \label{tab:READMEkeysEXP}
\end{table}

Because README.yaml contains all the essential information on each simulation, arbitrary analyses can be automatically performed over all the simualtions in the NMRlipids databank. In addition to the order parameters for each C-bonds and x-ray scattering form factors used in the quality evaluation, the NMRlipids databank contains the area per lipid and thickness calculated from all simulations in the databank. These results are stored in the same folders as the README.yaml files in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.

\subsection{Molecule and atom naming convention} \label{naming}

Because universal convention for lipid molecules and atoms therein has not been defined, the naming conventions vary between authors and force fields. To enable automatic analyses over large sets of simulation in the NMRlipids databank, we have defined unique naming conventions for lipid molecules and atoms. The abbreviations of molecule names used in the NMRlipids databank are listed in table~\ref{tab:abbreviations}. The unique atom names for each molecule and corresponding names in each simulation are defined using mapping files introduced in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).

Molecule and atom names in each simulation are connected to the unique naming convention with the COMPOSITION dictionary in the README.yaml file. Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary. The numbers of molecules in the system are then automatically calculated by the NMRlipids databank codes (AddData.py in Fig.~\ref{DatabankStructure}) and stored in the README.yaml together with other content of the COMPOSITION dictionary. This information can be then used to find each molecule and atom from each simulation in the analysis codes.

\begin{table}[h]
    \centering
    \begin{tabular}{c|c}
        Abbreviation & Molecule name \\
        \hline
        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
        CHOL & cholesterol \\
        DHMDMAB & dihexadecyldimethylammonium \\
        \hline
        POT & potassium ion \\
        SOD & sodium ion \\
        CLA & chloride ion \\
        CAL & calcium ion \\
        SOL & water \\
    \end{tabular}
    \caption{Abbreviations used in the databank}
    \label{tab:abbreviations}
\end{table}

\subsection{Quality evaluation}

>>>>

While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, researches are now searching new avenues by building models not only for individual molecules but whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.

@@

\subsection{Water diffusion anisotropy in membrane systems}
The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}. Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and theoretical methods, yet consensus 
%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}. MD simulations have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}. Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.

To this end, we first calculated the water permeability through membranes from all the simulations in the NMRlipids databank. Resulting permeabilities are shown as a function of temperature, membrane thickness, area per lipid, and acyl chain order in Figs.~\ref{fig:permeability} B-E. The values for water permeabilities from simulations vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}. To analyze how water permeability through membranes depend on bilayer properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range. These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots. As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B. Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order. Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence is observed for more loosely packed membranes. 
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
%, although weak decrease for the latter two may be visible.

@@

To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank. The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}. 
%The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
%As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
Simulation results are close to the experimental values with low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels. This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}. The water diffusion parallel to membrane surface increased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information. In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{tanner78,wasterby02}. The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}. Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} F and G.
This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig. \ref{fig:diffusionSI} A and C).

\subsection{Flip-flops}
Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}. Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}. Flip-flops have been previously studied mainly with coarse-grained simulations and free energy calculations due to time scale limitations in atomistic simulations~\cite{parisio16}, but direct analyses on cholesterol flip-flops in atomistic simulations have been also emerging in recent years~\cite{gu19,javanainen18,baral20}. However, systematic studies on how flip-flop rates depend on membrane properties are still challenging even with the state of the art computational resources. The NMRlipids databank can greatly advance such studies by giving an easy access to large amounts of data simultaneously.

@@

Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance. In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions. This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.

@@

Emerging applications of machine learning are increasing the impact of such databanks. For example, the existing Protein Databank (PDB)~\cite{??} containing experimentally determined protein structures with programmatic access has enabled the development of machine learning based tools building on the data collected in the databank over the years~\cite{??}. NMRlipids and other databanks with open programmatic access have potential to lead similar unforeseen applications in the future.

@@

\subsection{Structure of the databank}
Structure of the NMRlipids databank is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information. The databank locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}. Whenever a specific file is referred here, the file path within the NMRlipids databank repository is given. The scripts in the NMRlipids databank are mainly written in Python and many of them utilize the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.

The core content of the NMRlipids databank is stored in README.yaml files located at {\it /Data/Simulations} in the NMRlipids databank repository. Essential information for each simulation entry in the databank is stored in these files. These yaml files are both human and machine readable, and contain access to all information that are needed for further analyses of the simulations. The keys available in these files and their explanations are listed in table~\ref{tab:READMEkeys} in the supplementary information. To keep the databank light, the raw MD simulation data is stored in an external location which can be any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}. Whenever needed, the raw data can be downloaded through the links stored in the README.yaml files.

\subsection{Molecule and atom naming convention} \label{naming}
Molecule and atom names used in a simulation are often needed when analyzing the simulation trajectories. However, the naming conventions of lipid molecules and atoms therein vary between authors and force fields because universal naming convention has not been defined. Because such convention is necessary for automatic analyses over simulations, we have defined the unique naming conventions for lipid molecules and atoms in the NMRlipids databank. For each molecule we have defined an unique abbreviation which are listed in table~\ref{tab:abbreviations} in the supplementary information. On the other hand, unique names for each atom in a molecule are defined and connected to corresponding names in a simulation using mapping files. Mapping files are in yaml format where unique atom names are connected to force field specific names. Also residue names and fractions lipids (headgroup, glycerol bakcbone, and acyl chains) can be defined in the mapping file. For example, the mapping files used in the NMRlipids databank are located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids repository. For more details about mapping files, see~\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}). In practise, information about molecule and atom names are given in the COMPOSITION dictionary in README.yaml files, where force field specific molecule name corresponding the unique abbreviation (table~\ref{tab:abbreviations}) and the name of the mapping file are given.

%Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.

\subsection{Adding data into the NMRlipids databank}
In order to add a simulation entry into the NMRlipids databank, an initial info.yaml file containing user given information listed in table~\ref{tab:READMEkeys} needs to be created. For examples of these files, see {\it /Scripts/BuildDatabank/info\_files} in the NMRlipids databank repository. Using this file as an input, the {\it /Scripts/BuildDatabank/AddData.py} script then calculates the rest of the information to be stored in the README.yaml files, also listed in table~\ref{tab:READMEkeys}. The databank is open for submission of these info.yaml files via pull requests to the git repository. Large fraction of the current content is composed of the data uploaded to Zenodo in previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}, yet also other simulation data from the Zenodo repository are added.

\subsection{Experimental data}
Experimental data used in the quality evaluation, currently composed of C-H bond order parameters and x-ray scattering form factors, are stored in {\it /Data/experiments} in the NMRlipids databank repository. Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment. The keys and their descriptions for the experimental data are given in table~\ref{tab:READMEkeysEXP}. NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{??} and x-ray scattering data from Refs.~\citenum{??}. In addition, some previously unpublished NMR and x-ray scattering data are used. These are measured with previously established methods as described in the supplementary information.

\subsection{Calculation of C-H bond order parameters}
The C--H bond order parameters were calculated directly from the carbon and hydrogen positions using the definition
%The order parameters are defined as
\begin{equation}\label{OP}
S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
\end{equation}
where $\theta$ is the angle between the C--H bond and the membrane normal
%(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
and angular brackets denote the ensemble average, i.e., average over all sampled configurations of all lipids in a simulation. As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error bar~\cite{botan15}.

In practise, the analysis code loops over all README.yaml files (i.e., simulations in the NMRlipids databank), downloads the raw simulation to a local computer, uses the information about the atom and molecule naming conventions in README.yaml to determine the location of carbon and hydrogen atoms, and then calculates the order parameters from Eq.~\ref{OP}. The script to calculate all C-H bond order parameters from the NMRlipids databank is available at {\it /Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository. The resulting order parameters are stored for all simulations in files named \newline 
[{\it lipid\_name}]{\it OrderParameters.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.

\subsection{Calculation of x-ray scattering form factors}
X-ray scattering form factors were calculated with standard equation for symmetric lipid bilayers~\cite{ollila16}
\begin{equation}\label{FFsimpl}
F(q)=\int_{-D/2}^{D/2}\Delta \rho_e(z) \cos(zq_z) {\rm d}z,
\end{equation}
where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the z-direction. For the calculation of density profiles, atom coordinates were centred around the centre of mass of the POPC lipid molecules for every time frame, and a histogram of these centred positions was calculated with the bin width of $1/3$~Å. Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors. The script to calculate form factors for all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_FormFactors.py}.  The resulting form factors are stored for all simulations in files named {\it FormFactor.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.

\subsection{Calculation of membrane thickness, and area per lipids}
Lipid bilayer thicknesses in all simulations are calculated from the intersections of lipid and water electron densities using the script {\it Scripts/AnalyzeDatabank/calc\_thickness.py} at the NMRlipids databank repository. The resulting thicknesses are stored for all simulations in files named {\it thickness.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository. Lipid bilayer area per lipids in all simulations are calculated by dividing the simulation box area with the total number of molecules considered as lipids using the script {\it Scripts/AnalyzeDatabank/calcAPL.py} at the NMRlipids databank repository. The resulting area per lipids are stored for all simulations in files named {\it apl.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.

\subsection{Quality evaluation of C-H bond order parameters}
As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees. For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated. In practise, this is implemented by adding the path of the experimental data into the simulation README.yaml file using the {\it /Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids databank repository.

%\subsubsection{Conformational ensembles using C-H bond order parameters}
%
%The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.

The quality of each C--H bond order parameter is estimated by calculating the probability of a simulated value to locate within the error bars of the experimental value. Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer, 
$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom and the probability for a order parameter in simulation to locate within experimental error bars can be estimated from equation
\begin{equation}\label{probability}
  P = f \left( \frac{S_{\rm CH} - (S_{\rm exp}+\Delta S_{\rm exp})}{s/\sqrt{n}} \right) - f \left( \frac{S_{\rm CH} - (S_{\rm exp}-\Delta S_{\rm exp})}{s/\sqrt{n}} \right),
\end{equation}
where f(t) is the %first order 
Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C-H bond which equals the number of lipids in a simulation, $S_{\rm CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_{\rm CH}$ calculated over individual lipids, $S_{\rm exp}$ is the experimental value, and $\Delta S_{\rm exp}$ its error. The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}. Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution. However, the normal distribution gives probability values that are below the numerical accuracy of computers when simulation values are far from experiments. To avoid such numerical instabilities, we use the first order Student's t-distribution having slightly higher probabilities for values far away from the mean.
%Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids. 
On the other hand, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}. Such large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars. Therefore, the order parameters with simulation error bars above the experimental error 0.02 are not included in the quality evaluation.

%We estimate the probability for the simulation order parameter to locate within experimental
%error bars from equation
%\begin{equation}\label{probability}
%  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
%\end{equation}
%where T is the first order Student's t-distribution shifted to $\bar{X}$ and
%scaled with $s$. 
%where the probability is calculated from the normal distribution
%\begin{equation}\label{gaussian}
%    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
%    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
%\end{equation}
%$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations. The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge.

To streamline the comparison between simulations, we define the qualities of different fragments (headgroup, acyl chains or total lipid) within each lipid type in a simulation as
\begin{equation}
    P^{\rm frag}[{\rm lipid}] = \frac{\langle P[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
\end{equation}
where $P^{\rm frag}[{\rm lipid}]$ is the average of individual C-H bond order parameters qualities within the fragment, $p_{\rm frag}[{\rm lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule). The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
\begin{equation}
    P^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} P^{\rm frag}[{\rm lipid}],
\end{equation}
where $\chi_{\rm lipid}$ is the molar fraction of a lipid in the bilayer.

The quality evaluation of order parameters is implemented in {\it /Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository. The resulting qualities for each order parameters are stored in files named  [{\it lipid\_name}]{\it \_OrderParameters\_quality.json}, for individual lipids in files named [{\it lipid\_name}]{\it \_FragmentQuality.json}, and for overall quality for fragments in files named {\it system\_quality.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.

\subsection{Quality evaluation of x-ray scattering form factors}
%While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
Because experiments give form factors only in relative intensity scale, they should scaled before comparing with the simulation data. Here we use the scaling coefficient for experimental intensities defined in the SIMtoEXP program~\cite{kucerka10}
\begin{equation}
    k_e = \frac{\sum_{i=1}^{N_q} \frac{|F_s(q_i)||F_e(q_i)|}{(\Delta F_e(q_i))^2}}{\sum_{i=1}^{N_q} \frac{|F_e(q_i)|^2}{(\Delta F_e(q_i))^2}},
\end{equation}
%The quality of each form factor was then calculated from the equation
%\begin{equation}
%    \chi^2 = \frac{\sqrt{\sum_{i=1}^{N_q}(|F_s(q_i)|-k_e|F_e(q_i)|)^2/(\Delta F_e(q_i))^2}}{\sqrt{N_q-1}},
%\end{equation}
where $F_s(q)$ and $F_e(q)$ are form factors from a simulation and experiment, respectively, $\Delta F_e(q)$ is the error of the experimental form factor, and summation goes over the experimentally available $N_q$ points.
